---
title: "Time-period specific HRs"
output: html_document
---

```{r, warning=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

## Specify whether to visualise unadjusted, partially adjusted, or fully adjusted models
model_type = "partial" # select from: partial or full

## Import libraries
library('here')
library('tidyverse')
library('cowplot')

## Import data
data <- read_csv(here("all_cox_data.csv")) %>%
  mutate(ref = Ethnicity=="White")
data$Period = factor(data$Period, levels = unique(data$Period))
data$Redaction_flag = NA
data$Redaction_flag[data$Denominator=="redacted"] = "[R]"

## Read in selected HRs and CIs
if (model_type=="full") {
  data$HR = data$F_adj_hr
  data$lowerCI = data$F_adj_lci
  data$upperCI = data$F_adj_uci
  data$HR_ratio = data$F_HRR_vs_pre
} else {
  data$HR = data$P_adj_hr
  data$lowerCI = data$P_adj_lci
  data$upperCI = data$P_adj_uci
  data$HR_ratio = data$P_HRR_vs_pre
}

## Set factor levels
data$Ethnicity = factor(data$Ethnicity, levels = rev(unique(data$Ethnicity)))
data$Outcome = factor(data$Outcome, levels = unique(data$Outcome))
```

### Visualisation parameters
* Pre-pandemic rates used as reference group throughout    
* Selected HR models: `r model_type`
* [R] used in HR and IR ratio plots to indicate redacted comparisons    
<br><br>



### Absolute and relative risk over time (pre vs pandemic)
```{r}
## Function to generate grid of IR/HR plots for pre vs post pandemic
pre_post_grid_plot = function(selected_group) {
  
  ## Select subset of data corresponding to prepandemic and pandemic periods
  d1 = subset(data, Period=="Pre" | Period=="Pandemic") %>% 
      ## Arrange with pre-pandemic period last to make sure this point is on top (and therefore visible)
      arrange(rev(Period)) 
  
  ## Select subset of data corresponding to specified outcome group
  if (selected_group!="All") { d1 = subset(d1, Group==selected_group) }

  ## Create IR plot
  d1$facet = "IRs over time"
  p1 = ggplot(d1, aes(y = Rate, x = Ethnicity, colour = Period)) + 
      geom_point(size = 4, alpha=0.8) +
      coord_flip() + facet_grid(Outcome~facet) + 
      theme_bw() + theme(strip.background = element_blank()) + 
      scale_colour_manual(values = c("Pre" = "#993404", "Pandemic" = "#045a8d")) +
      ylab("Rate per 100,000 person-months") + xlab("") +
      theme(axis.title = element_text(size=12), axis.text = element_text(size=12), 
           strip.text.x = element_text(size=12), strip.text.y = element_blank(), 
           legend.position = "left", legend.title = element_text(size=12), legend.text = element_text(size=12), 
           legend.direction = "vertical")
  
 ## Create IR plot
  d1$facet = "IRs over time"
  p1 = ggplot(d1, aes(y = Rate, x = Ethnicity, colour = Period)) + 
      geom_point(size = 4, alpha=0.8) +
      coord_flip() + facet_grid(Outcome~facet) + 
      theme_bw() + theme(strip.background = element_blank()) + 
      scale_colour_manual(values = c("Pre" = "#993404", "Pandemic" = "#045a8d")) +
      ylab("Rate per 100,000 person-months") + xlab("") +
      theme(axis.title = element_text(size=12), axis.text = element_text(size=12), 
           strip.text.x = element_text(size=12), strip.text.y = element_blank(), 
           legend.position = "left", legend.title = element_text(size=12), legend.text = element_text(size=12), 
           legend.direction = "vertical")
  
  ## Create IR ratio plot
  d1$facet = "Ratio vs\npre-pandemic"
  p2 = ggplot(d1, aes(y = IRR_vs_pre, x = Ethnicity)) + 
      geom_bar(stat="identity", fill="#045a8d") +
      geom_text(aes(y = 0.3, label = Redaction_flag), colour = "#045a8d", show.legend = FALSE) +
      coord_flip() + facet_grid(Outcome~facet) + 
      theme_bw() + theme(strip.background = element_blank()) + 
      geom_hline(yintercept=1, linetype="dotted") + 
      scale_y_continuous(trans="log10", limits=c(0.3,3), breaks=c(0.5,1,2)) +
      ylab("IR ratio") + xlab("") +
      theme(axis.title = element_text(size=12), axis.text = element_text(size=12), 
             strip.text.x = element_text(size=12), strip.text.y = element_text(size=12), 
            legend.position = "none", axis.text.y=element_blank())
  
  ## Create HR plot
  d1$facet = "a"
  p3 = ggplot(d1, aes(y = HR, x = Ethnicity, colour = Period, shape = ref)) + 
      geom_point(size = 4, alpha=0.8) +
      geom_errorbar(aes(ymin=as.numeric(lowerCI), ymax=as.numeric(upperCI)), width=0) + 
      coord_flip() + facet_grid(Outcome~facet) + 
      theme_bw() + theme(strip.background = element_blank()) + 
      geom_hline(yintercept=1, linetype="dotted") + 
      scale_colour_manual(values = c("Pre" = "#993404", "Pandemic" = "#045a8d")) +
      scale_y_continuous(trans="log10") +
      scale_shape_manual(values=c(19, 1)) + guides(shape = "none") +
      ylab("HR vs White") + xlab("") +
      theme(axis.title = element_text(size=12), axis.text = element_text(size=12), 
            strip.text.x = element_text(size=12), strip.text.y = element_blank(), 
           legend.position = "bottom", legend.title = element_text(size=12), legend.text = element_text(size=12), 
           legend.direction = "vertical")
  
  ## Create HR ratio plot
  d1$facet = "b"
  p4 = ggplot(d1, aes(y = HR_ratio, x = Ethnicity)) + 
      geom_bar(stat="identity", fill="#045a8d") +
      geom_text(aes(y = 0.3, label = Redaction_flag), colour = "#045a8d", show.legend = FALSE) +
      coord_flip() + facet_grid(Outcome~facet) +    
      theme_bw() + theme(strip.background = element_blank()) + 
      geom_hline(yintercept=1, linetype="dotted") + 
      scale_y_continuous(trans="log10", limits=c(0.3,3), breaks=c(0.5,1,2)) +
      ylab("HR vs pre-pandemic") + xlab("") +
      theme(axis.title = element_text(size=12), axis.text = element_text(size=12), 
             strip.text.x = element_text(size=12), strip.text.y = element_text(size=10), 
            legend.position = "left", legend.title = element_text(size=12), legend.text = element_text(size=12), 
           legend.direction = "vertical")

  ## Render grid of plots
  plot_grid(p3, p4, ncol=4, rel_widths=c(1,0.5,1,0.5), align="h", axis="tb")
}
```


### Absolute risk over time (wave/easing)
```{r}
## Function to generate grid of IR or HR plots for pre vs pandemic, broken down by wave/easing
pre_wave_easing_grid_plot = function(selected_group, outcome=c("IR", "HR", "IRD")) {
  
  ## Select subset of data corresponding to pandemic sub-periods
  d1 = subset(data, Period!="Pandemic") %>%
    ## Arrange with pre-pandemic period last to make sure this point is on top (and therefore visible)
    arrange(rev(Period))
  
  ## Select subset of data corresponding to specified outcome group
  if (selected_group!="All") { d1 = subset(d1, Group==selected_group) }
  
  ## Hex palette for wave/easing colours
  pal = c("#023858", "#045a8d", "#0570b0", "#3690c0", "#74a9cf", "#a6bddb")

  ## Specify plot labels and variables depending on whether IR or HR selected
  if(outcome=="IR") {
    d1$facet = "IRs over time"
    axis_lab_primary = "Rate per 100,000 p-w"
    d1$selected_y_primary = d1$Rate
    axis_lab_secondary = "IR ratio"
    d1$selected_y_secondary = d1$IRR_vs_pre
    d1$ref = TRUE 
  } else if(outcome=="IRD") {
    d1$facet = "IRs over time"
    axis_lab_primary = "Rate per 100,000 p-w"
    d1$selected_y_primary = d1$Rate
    axis_lab_secondary = "IR difference"
     d1$selected_y_secondary = d1$IRD_vs_pre
  } else {
    d1$facet = ""
    axis_lab_primary = "HR vs White"
    d1$selected_y_primary = d1$HR
    axis_lab_secondary = "HR ratio"
    d1$selected_y_secondary = d1$HR_ratio
  }
  
  ## Create IR/HR plot
  p1 = ggplot(d1, aes(y = selected_y_primary, x = Ethnicity, colour = Period, shape = ref)) + 
      geom_point(size = 4, alpha=0.8) +
      coord_flip() + facet_grid(Outcome~facet) + 
      theme_bw() + theme(strip.background = element_blank()) + 
      scale_colour_manual(values = c("#993404", pal[1:6])) +
      ylab(axis_lab_primary) + xlab("") +
      scale_shape_manual(values=c(19, 1)) + guides(shape = "none") +
      theme(axis.title = element_text(size=12), axis.text = element_text(size=12), 
           strip.text.x = element_text(size=12), strip.text.y = element_blank(), 
           legend.position = "top", legend.title = element_blank(), legend.text = element_text(size=8), 
           legend.direction = "horizontal")
  if (outcome == "HR") {
    p1 = p1 + geom_hline(yintercept=1, linetype="dotted") 
  }
  
  ## Function to add a consistent set of plotting parameters to each of the ggplots below
  if (outcome != "IRD") {
    gg_extras = function(base_plot) {
      base_plot +  coord_flip() + facet_grid(Outcome~facet) + 
        theme_bw() + theme(strip.background = element_blank()) + 
        geom_hline(yintercept=1, linetype="dotted") + 
        scale_y_continuous(trans="log10", limits=c(0.3,3), breaks=c(0.5,1,2)) +
        ylab(axis_lab_secondary) + xlab("") +
        theme(axis.title = element_text(size=12), axis.text = element_text(size=12), 
               strip.text.x = element_text(size=12), strip.text.y = element_text(size=12), 
              legend.position = "none", axis.text.y=element_blank())
    }
  }
  else {
    gg_extras = function(base_plot) {
      base_plot +  coord_flip() + facet_grid(Outcome~facet) + 
        theme_bw() + theme(strip.background = element_blank()) + 
        geom_hline(yintercept=1, linetype="dotted") + 
        scale_y_continuous(limits=c(-5,5), breaks=c(-4,-2,0,2,4)) +
        ylab(axis_lab_secondary) + xlab("") +
        theme(axis.title = element_text(size=12), axis.text = element_text(size=12), 
               strip.text.x = element_text(size=12), strip.text.y = element_text(size=12), 
              legend.position = "none", axis.text.y=element_blank())
    }
  }
  
  ## Wave 1 ratio plot
  d1$facet = "Wave 1 vs\npre"
  p2 = gg_extras(ggplot(subset(d1, Period=="Pre" | Period=="Wave 1"), aes(y = selected_y_secondary, x = Ethnicity)) + 
      geom_text(aes(y = 0.5, label = Redaction_flag), colour = pal[1], show.legend = FALSE) +
      geom_bar(stat="identity", fill=pal[1])) + theme(strip.text.y = element_blank()) 
  
  ## Easing 1 ratio plot
  d1$facet = "Easing 1 vs\npre"
  p3 = gg_extras(ggplot(subset(d1, Period=="Pre" | Period=="Easing 1"), aes(y = selected_y_secondary, x = Ethnicity)) + 
      geom_text(aes(y = 0.5, label = Redaction_flag), colour = pal[1], show.legend = FALSE) +
      geom_bar(stat="identity", fill=pal[2])) + theme(strip.text.y = element_blank())
  
   ## Wave 2 ratio plot
  d1$facet = "Wave 2 vs\npre"
  p4 = gg_extras(ggplot(subset(d1, Period=="Pre" | Period=="Wave 2"), aes(y = selected_y_secondary, x = Ethnicity)) + 
      geom_text(aes(y = 0.5, label = Redaction_flag), colour = pal[1], show.legend = FALSE) +
      geom_bar(stat="identity", fill=pal[3])) + theme(strip.text.y = element_blank())
  
  ## Easing 2 ratio plot
  d1$facet = "Easing 2 vs\npre"
  p5 = gg_extras(ggplot(subset(d1, Period=="Pre" | Period=="Easing 2"), aes(y = selected_y_secondary, x = Ethnicity)) + 
      geom_text(aes(y = 0.5, label = Redaction_flag), colour = pal[1], show.legend = FALSE) +
      geom_bar(stat="identity", fill=pal[4])) + theme(strip.text.y = element_blank())
  
   ## Wave 3 ratio plot
  d1$facet = "Wave 3 vs\npre"
  p6 = gg_extras(ggplot(subset(d1, Period=="Pre" | Period=="Wave 3"), aes(y = selected_y_secondary, x = Ethnicity)) + 
      geom_text(aes(y = 0.5, label = Redaction_flag), colour = pal[1], show.legend = FALSE) +
      geom_bar(stat="identity", fill=pal[5])) + theme(strip.text.y = element_blank())
  
  ## Easing 3 ratio plot
  d1$facet = "Easing 3 vs\npre"
  p7 = gg_extras(ggplot(subset(d1, Period=="Pre" | Period=="Easing 3"), aes(y = selected_y_secondary, x = Ethnicity)) + 
      geom_text(aes(y = 0.5, label = Redaction_flag), colour = pal[1], show.legend = FALSE) +
      geom_bar(stat="identity", fill=pal[6]))
  
  ## Render grid of plots
  plot_grid(p1, p2, p3, p4, p5, p6, p7, ncol=7, rel_widths=c(1,rep(0.5,5),0.6), align="h", axis="tb")
}
```


#### Cardiovascular stroke & vte
```{r, fig.width=16, fig.height=5}
## Run function for each outcome group
## N.B. fig.height needs to be adjusted in the specifications for each chunk due to the different number of outcomes in each group
pre_post_grid_plot("Cardiovascular")
```
#### Cardiovascular - stroke & VTE
```{r, fig.width=14, fig.height=7}
## Run function for each outcome group
## N.B. fig.height needs to be adjusted in the specifications for each chunk due to the different number of outcomes in each group
pre_wave_easing_grid_plot("Cardiovascular", "IR")
```
#### Cardiovascular - stroke & VTE
```{r, fig.width=14, fig.height=7}
## Run function for each outcome group
## N.B. fig.height needs to be adjusted in the specifications for each chunk due to the different number of outcomes in each group
pre_wave_easing_grid_plot("Cardiovascular", "HR")
```

#### Cardiovascular MI & heart failure
```{r, fig.width=16, fig.height=5}
## Run function for each outcome group
## N.B. fig.height needs to be adjusted in the specifications for each chunk due to the different number of outcomes in each group
pre_post_grid_plot("Cardiovascular_hfmi")
```
#### Cardiovascular - heart failure & MI
```{r, fig.width=14, fig.height=7}
## Run function for each outcome group
## N.B. fig.height needs to be adjusted in the specifications for each chunk due to the different number of outcomes in each group
pre_wave_easing_grid_plot("Cardiovascular_hfmi", "IR")
```
#### Cardiovascular - heart failure & MI
```{r, fig.width=14, fig.height=7}
## Run function for each outcome group
## N.B. fig.height needs to be adjusted in the specifications for each chunk due to the different number of outcomes in each group
pre_wave_easing_grid_plot("Cardiovascular_hfmi", "HR")
```
<br><br>
#### Diabetes
```{r, fig.width=16, fig.height=5}
pre_post_grid_plot("Diabetes")
```
#### Diabetes
```{r, fig.width=12, fig.height=10}
pre_wave_easing_grid_plot("Diabetes", "IR")
```
#### Diabetes
```{r, fig.width=12, fig.height=10}
pre_wave_easing_grid_plot("Diabetes", "HR")
```
<br><br>
#### Respiratory
```{r, fig.width=16, fig.height=5}
pre_post_grid_plot("Respiratory")
```
#### Respiratory
```{r, fig.width=12, fig.height=7}
pre_wave_easing_grid_plot("Respiratory", "IR")
```
#### Respiratory
```{r, fig.width=12, fig.height=7}
pre_wave_easing_grid_plot("Respiratory", "HR")
```
<br><br>
#### Mental Health
```{r, fig.width=16, fig.height=5}
pre_post_grid_plot("Mental Health")
```
#### Mental Health
```{r, fig.width=12, fig.height=10}
pre_wave_easing_grid_plot("Mental Health", "IR")
```
#### Mental Health
```{r, fig.width=12, fig.height=7}
pre_wave_easing_grid_plot("Mental Health", "HR")
```
<br><br>
#### All
```{r, fig.width=16, fig.height=10}
pre_post_grid_plot("All")
```
#### All
```{r, fig.width=16, fig.height=20}
pre_wave_easing_grid_plot("All", "IR")
```
#### All
```{r, fig.width=16, fig.height=20}
pre_wave_easing_grid_plot("All", "HR")
```




#### Session info
```{r}
print(sessionInfo())
```

